---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(here)
```


## Stroke activity by local authority area.

```{r}
stroke_activitybyca <- read_csv("raw_data/stroke_activitybyca.csv") %>%
  clean_names()

head(stroke_activitybyca)
```

```{r}
stroke_activitybyca %>%
  distinct(ca)

```

subsetting data

```{r}
stroke_ca_2 <- stroke_activitybyca %>%
  select(financial_year, ca, admission_type  ,age_group, sex, diagnosis, number_of_discharges, crude_rate)
```


## Stroke activity by health board region.

```{r}
stroke_activitybyhbr <- read_csv("raw_data/stroke_activitybyhbr.csv") %>%
  clean_names()

head(stroke_activitybyhbr)
```

```{r}
stroke_activitybyhbr %>%
  distinct(hbr)
```
```{r}
stroke_hbr_2 <- stroke_activitybyhbr %>%
  select(financial_year, hbr, admission_type, age_group, sex, diagnosis, number_of_discharges, crude_rate)
```



Subsetting stroke mortality by local authority area
```{r}
stroke_mortalitybyca <- read_csv("raw_data/stroke_mortalitybyca.csv") %>%
  clean_names()

head(stroke_mortalitybyca)
```

```{r}
stroke_mortalitybyca_2 <- stroke_mortalitybyca %>%
  select(year, ca, age_group, sex, diagnosis, number_of_deaths, crude_rate)
```



Subsetting stroke mortality by health board region.
```{r}
stroke_mortalitybyhbr <- read_csv("raw_data/stroke_mortalitybyhbr.csv") %>%
  clean_names()

head(stroke_mortalitybyhbr)
```

```{r}
stroke_mortalitybyhbr_2 <- stroke_activitybyhbr %>%
  select(financial_year, hbr, age_group, sex, diagnosis, number_of_discharges, crude_rate)
```



```{r}
la_data_zones <- read_csv("raw_data/data_zones.csv")
```

```{r}
data_zones <- read_csv("~/public_health_dashboard/raw_data/data_zones.csv")

la_zones_clean <- data_zones %>%
  select(LA_Code, LA_Name) %>%
  distinct()
  
```

```{r}
hbr_zones_clean <- data_zones %>%
  select(HB_Code, HB_Name) %>%
  distinct()
  
```



```{r}
comb_stroke_activityca <- left_join(stroke_ca_2, la_zones_clean, by = c("ca" ="LA_Code"))
```

```{r}
comb_stroke_activityhbr <- left_join(stroke_hbr_2, hbr_zones_clean, by = c("hbr" ="HB_Code"))
```

```{r}
comb_stroke_mortalityca <- left_join(stroke_mortalitybyca_2, la_zones_clean, by = c("ca" ="LA_Code"))
```

```{r}
comb_stroke_mortalityhbr <- left_join(stroke_mortalitybyhbr_2, hbr_zones_clean, by = c("hbr" ="HB_Code"))
```


```{r}
library(ggplot2)

```

```{r}
ggplot(comb_stroke_activityca) +
  
```

```{r}
common_stroke <- comb_stroke_activityca %>%
  filter( ca == "S92000003") %>%
  filter(age_group %in% c("All")) %>%
  filter(admission_type %in% c("All")) %>%
  filter(sex %in% c("All"))
           
  
```


```{r}
ggplot(common_stroke) +
  aes(x = diagnosis, y = number_of_discharges) +
  geom_col() 
```

```{r}
common_stroke_2 <- common_stroke %>%
  group_by(diagnosis) %>%
  mutate(number_of_cases = sum(number_of_discharges)) %>%
  filter(financial_year == "2009/10")
```


```{r}
ggplot(common_stroke_2) +
  aes(x = diagnosis, y = number_of_cases) +
  geom_col(fill = "blue") +
  ggtitle("Prevalence of Cases of Types of Stroke Across All of Scotland for the years 2009 to 2019") +
  xlab("Diagnosis") +
  ylab("Number of Cases") +
  theme(axis.text.x = element_text(angle = 330))+
  theme(plot.title = element_text(hjust = 0.5))
```

Differences between demographics
```{r}
common_stroke_dem <- comb_stroke_activityca %>%
  filter( ca == "S92000003") %>%
  filter(age_group %in% c("All")) %>%
  filter(admission_type %in% c("All")) %>%
  filter(sex %in% c("Male", "Female"))  %>%
  group_by(diagnosis, sex) %>%
  mutate(number_of_cases = sum(number_of_discharges)) %>%
  filter(financial_year == "2009/10")
          
```


```{r}
ggplot(common_stroke_dem) +
  aes(x = diagnosis, y = number_of_cases, fill = sex) +
  geom_col() +
  labs(fill= "Gender")
```




```{r}
common_stroke_year <- comb_stroke_activityca %>%
  filter( ca == "S92000003") %>%
  filter(age_group %in% c("All")) %>%
  filter(admission_type %in% c("All")) %>%
  filter(sex %in% c("Male", "Female"))  %>%
  group_by(diagnosis, sex)
```


```{r}
ggplot(common_stroke_year) +
  aes(x=diagnosis, y=number_of_discharges, fill=sex)+
  geom_col() +
  facet_wrap(~financial_year) +
  ggtitle("Number of Cases of Stroke by Type in Males and Females per year") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Number of Cases") +
  labs(fill= "Gender")
  
```
ggplot(count_day) +
  aes(x=start_day,y=counts, colour=start_month)+
  geom_point() +
  geom_line() +
  facet_wrap(~start_month) +
  ggtitle("Number of Citi Bike Hires Per Month") +
  xlab("Day of Month") +
  ylab("Number of Bike Hires")+
  theme_minimal()

number of cases per year
```{r}
ggplot(common_stroke) +
  aes(x = diagnosis, y = number_of_discharges) +
  geom_col(fill = "blue") +
  facet_wrap(~financial_year) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
common_stroke_age <- comb_stroke_activityca %>%
  filter( ca == "S92000003") %>%
  filter(admission_type %in% c("All")) %>%
  filter(sex %in% c("All")) %>%
  group_by(diagnosis, age_group)
```

```{r}
ggplot(common_stroke_age) +
  aes(x = diagnosis, y = crude_rate) +
  geom_col(fill = "blue") +
  facet_wrap(~age_group) +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Number of Cases of Each Type of Stroke by Age Group for the years 2009 to 2019") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Stroke Type") +
  ylab("Number of Cases")
```



```{r}
test_1 <- common_stroke_age %>%
  filter(diagnosis == "Cerebrovascular Disease" & age_group == "All") 
  
test2 <- separate(test_1, financial_year, into = c("year"), sep = "/")
```

separate(meteorites_clean_b, geo_location, into = c("latitude", "longitude"),sep=", ")


```{r}


test2 <- transform(test2, year = as.numeric(year))

```



```{r}
common_stroke_area <- separate(common_stroke, financial_year, into = c("year"), sep = "/")
  
  
common_stroke_area_2 <- transform(common_stroke_area, year = as.numeric(year))

  
```

```{r}
ggplot(common_stroke_area_2) +
  aes(x = year, y = number_of_discharges, fill = diagnosis) +
  geom_area() +
  ggtitle("Number of Cases of Each Type of Stroke per Year") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Year") +
  ylab("Number of Cases") +
  labs(fill= "Type of Stroke") 
  
```

```{r}
library(sf)
library(leaflet)
```



```{r}

map <- st_read(here("~/final_project_melanie_jayasinghe/clean_data/shape_data/pub_las.shp")) %>%
  st_simplify(dTolerance = 1000) %>%
  st_transform("+proj=longlat + datum=WGS84")

```

```{r}
plot(map)
```
```{r}
hbr_map <- st_read(here("~/final_project_melanie_jayasinghe/clean_data/shape_data_2/SG_NHS_HealthBoards_2019.shp")) %>%
  st_simplify(dTolerance = 1000) %>%
  st_transform("+proj=longlat + datum=WGS84")

```

```{r}
plot(hbr_map)
```


```{r}
hbr_map_data <- comb_stroke_activityhbr %>%
  filter(admission_type == "All", age_group == "All", sex == "All")
```

```{r}
# Filter data for leaflet plot
stroke_type_map <- hbr_map_data %>%
  group_by(hbr) %>%
  # Join with shape data
  left_join(hbr_map, by = c("hbr" = "HBCode")) %>%
  # Convert to shape data from data frame
  st_as_sf()
```





